---
title: "ADA MID"
author: "yoseph"
date: "10/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

a) importing the dataset
```{r}
library(haven)
library(readr)
library(dplyr)
library(tidyverse)
df = read.csv('https://github.com/kijohnson/Data/blob/main/titanic.csv?raw=true')
```

b) determining the median age for by passenger class
```{r}
df %>%
    group_by(Pclass) %>%
    dplyr::summarize(median_age = median(Age, na.rm=TRUE))


```
interpretation of the median 
The median age for class 1 passanger is 38.5 years old
The median age for class 2 passangers is 29 years old
The median age for class 3 passangers is 24 years old

c) a plot that shows medians, spread of data, and outliers for the continuous variable age according to passenger class
### to chech the median on the graph 
##1
```{r}
library (tidyverse)
ggplot(data = df) + 
  stat_summary(
    mapping = aes(x = Pclass, y = Age),
    fun.min = min,
    fun.max = max,
    fun = median)

```
Or 
##2 checking the outliers, median, and data distribution at the same time

```{r}
a = ggplot(df, aes(x = factor(Pclass), y = Age, fill = factor(Pclass))) + 
  geom_boxplot() + 
  stat_summary(
    aes(label = round(stat(y), 1)),
    geom = "text", 
    fun = function(y) { o <- boxplot.stats(y)$out; if(length(o) == 0) NA else o },
    hjust = -1, size = 2
  )
a

```

##Printing the graph and the outlier table to 
```{r}
png(filename="v.jpg")
plot(a)
dev.off()
outliers = boxplot(df$Age, plot=FALSE)$out
outliers
``````
```{r}
df %>%
    group_by(Pclass) %>%
    dplyr::summarize(sum(df$Sex == "male")/length(df$Sex), sum(df$Sex == "female")/length(df$Sex))

```


